Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-08-12 ms.tex
Updated Test 9 opt. thin figures.  Wrote description of Test 9.

\begin{figure}[t]
  \epsscale{0.5}
  \plotone{fig/test9thick_rho}
  \plotone{fig/test9thick_temp}
  \plotone{fig/test9thick_efrac}
  \caption{\label{fig:test9_thick} Test 9. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).}
  \epsscale{1}
\end{figure}

\begin{figure}[t]
  \epsscale{0.5}
  \plotone{fig/test9thin_rho}
  \plotone{fig/test9thin_temp}
  \plotone{fig/test9thin_efrac}
  \caption{\label{fig:test9_thin} Test 9. (Irradiated Rayleigh-Taylor
    instability; optically thin case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).}
  \epsscale{1}
\end{figure}

Here we combine the classic case of a Rayleigh-Taylor instability and
an expanding \ion{H}{2} region.  The Rayleigh-Taylor instability
occurs when a dense fluid is being supported by a lighter fluid,
initially in hydrostatic equilibrium, in the presence of a constant
acceleration field.  This classic test alone evaluates how subsonic
perturbations evolve.  We consider the case of a single-mode
perturbation.  We let the system evolve without any radiation until
the perturbation grows considerably and then turn on the radiation
source.

@@

The optically-thin case is set up similarly but with three
changes---(1) a density constrast of 10, (2) a luminosity of $10^{43}$
ph s$^{-1}$, and (3) the source is born at 6.5 Myr. The time units are
decreased to 200 kyr so that $(T_0, T_1) = (1.8 \times 10^3, 1.8
\times 10^4)$ K.  Note that in code units, pressure is unchanged.  We
adjust the physical unit scaling, so the light fluid has $T > 10^4$ K,
$x_e \sim 1$, and thus optically-thin.  A possible physical analogue
could be a radiation source heating and rarefying the medium below.

The $x$ and $y$-boundaries are periodic, and the $z$-boundaries are
reflecting.  The radiation source is turned on at $t = 7.5$ of
evolution and placed at the center of lower $z$-boundary face.  The
periodic boundaries will cause features that are non-physical,
considering a more realistic plane-parallel case.  Nevertheless, these
tests provide a good check on a radiation hydrodynamics solver.

@@

Next we construct a problem that is similar to the Rayleigh-Taylor
test problem (Test 9) in nature but is spherically symmetric.  We use
the same physical setup as Test 5 with a uniform medium and the
radiation source at the origin.  It has a luminosity of $5 \times
10^{48}$ ph s$^{-1}$ and a mono-chromatic spectrum of 14.6 eV.  We
choose this energy so the ionization front is relatively sharp but
still heats the gas to $10^4$ K.  The ambient medium has a density of
$10^{-3} \cubecm$ and an initial temperature of 100 K.  A point source
of gravity also resides at the origin.  To have an equal gravitational
acceleration $g(r) = g_0 / r^3$ at the Str\"{o}mgren radius $R_s =
5.4$ kpc ($r = 0.36$) as Test 9, we set the strength of the point
source $g_0 = 4.67 \times 10^{-3}$.  The gravity point source has the
same effect on any pertubations as the constant gravitational force in
Rayleigh-Taylor instability test.

>>>>

\begin{figure}[t]
  \epsscale{0.5}
  \plotone{fig/test9thick_rho}
  \plotone{fig/test9thick_temp}
  \plotone{fig/test9thick_efrac}
  \caption{\label{fig:test9_thick} Test 9. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).  The
    source turns on at $t=0$.}
  \epsscale{1}
\end{figure}

\begin{figure}[t]
  \epsscale{0.5}
  \plotone{fig/test9thin_rho}
  \plotone{fig/test9thin_temp}
  \plotone{fig/test9thin_efrac}
  \caption{\label{fig:test9_thin} Same as Figure \ref{fig:test9_thick}
    but for the optically thin case.}
  \epsscale{1}
\end{figure}

Here we combine the classic case of a Rayleigh-Taylor instability and
an expanding \ion{H}{2} region.  The Rayleigh-Taylor instability
occurs when a dense fluid is being supported by a lighter fluid,
initially in hydrostatic equilibrium, in the presence of a constant
acceleration field.  This classic test alone evaluates how subsonic
perturbations evolve.  We consider the case of a single-mode
perturbation.  We let the system evolve without any radiation until
the perturbation grows considerably and then turn on the radiation
source.  These tests demonstrate that Enzo+Moray can follow a highly
dynamic system and resolve fine density structures.

@@

The optically-thin case is set up similarly but with three
changes---(1) a density constrast of 10, (2) a luminosity of $10^{43}$
ph s$^{-1}$, and (3) the source is born at 6.5 Myr.  The time units
are decreased to 200 kyr so that $(T_0, T_1) = (1.8 \times 10^3, 1.8
\times 10^4)$ K.  Note that in code units, pressure is unchanged.  We
adjust the physical unit scaling, so the light fluid has $T > 10^4$ K,
$x_e \sim 1$, and thus optically-thin.  Furthermore, the ionization
front remains R-type before interacting with the instability.  A
possible physical analogue could be a radiation source heating and
rarefying the medium below.

The $x$ and $y$-boundaries are periodic, and the $z$-boundaries are
reflecting.  The radiation source is placed at the center of lower
$z$-boundary face.  The periodic boundaries will cause features that
are non-physical, considering a more realistic plane-parallel case.
Nevertheless, these tests provide a good check on a radiation
hydrodynamics solver.  We show the evolution of the density,
temperature, and ionized fraction of the optically thick and optically
thin cases in Figures \ref{fig:test9_thick} and \ref{fig:test9_thin}.
The initial state of the Rayleigh-Taylor instability is shown in the
left panels.

In the optically thick case, a D-type front is created, which is
clearly illustrated by the spherical density enhancement at 0.02 Myr.
The shock then passes through the instability at $\sim$0.25 Myr and
reflects off the upper $z$-boundary.  This and complex shock
reflections create a Richtmeyer-Meshkov instability, driving a chaotic
jet-like structure downwards.  The radiation source photo-evaporates
the outer parts of this structure.  The interaction between the dense
cool ``jet'' and the hot medium further drives instabilities along the
surface, which can be seen when comparing $t = 0.59$ Myr and $t =
0.91$ Myr slices.  At the latter time, the jet cannot reach the bottom
of the domain before being photo-evaporated.  Eventually this
structure is completely destroyed, leaving behind a turbulent medium
between the hot and cold regions.

The optically thin problem is less violent than the optically thick
case because the R-type front does not interact with the initial
instability as strongly.  The radiation source provides further
buoyance in the already $T=10^4$ K gas.  The gas first to be ionized
and photo-evaporates is the outer regions of the instability.  The
enhanced heating also drives the upper regions of the instability,
making the top interface turbulent.  It then reflects off the upper
$z$-boundary and creates a warm $T = 5 \times 10^3$ K, partially
ionized ($x_e \sim 10^{-2}$), turbulent medium, seen in the slices $t
\ge 0.67$ Myr.  The slices of electron fraction also show that the
dense gas is optically thick.

@@

Next we construct a problem that is similar to the Rayleigh-Taylor
test problem (Test 9) in nature but is spherically symmetric.  We use
the same physical setup as Test 5 with a uniform medium and the
radiation source at the origin.  It has a luminosity of $5 \times
10^{48}$ ph s$^{-1}$ and a mono-chromatic spectrum of 14.6 eV.  We
choose this energy so the ionization front is relatively sharp but
still heats the gas to $10^4$ K.  The ambient medium has a density of
$10^{-3}~\cubecm$ and an initial temperature of 100 K.  A point source
of gravity also resides at the origin.  To have an equal gravitational
acceleration $g(r) = g_0 / r^3$ at the Str\"{o}mgren radius $R_s =
5.4$ kpc ($r = 0.36$) as Test 9, we set the strength of the point
source $g_0 = 4.67 \times 10^{-3}$.  The gravity point source has the
same effect on any pertubations as the constant gravitational force in
Rayleigh-Taylor instability test.
